\textbf{Discovery of a disrupting open cluster in the Galactic halo:\\ a recent star formation event in the leading arm of the Magellanic stream}

### 2018-11-04 ms.tex
twocolumn

% TODO
% - Sand, Seth, et al. Virgo: blue stars with nothing else
% - Keywords after abstract
% - Split cluster into 2? Price-Whelan 1a and 1b?

\documentclass[modern]{aastex62}
\usepackage{amsmath}
\graphicspath{{./}{figures/}}

@@

We report the discovery of a young $\tau \sim \clage$, low-mass $M \sim 1200~\msun$, metal-poor $[\textrm{Fe}/\textrm{H}] \sim \clfeh$ stellar association in the Milky Way halo, \clustername.
At a Galactocentric position $(R, z) \sim (23, 15)~\textrm{kpc}$, the association is (on the sky) near the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the Magellanic stream, $D \approx \cldist$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way that do not account for ram pressure and gas interaction with MW disk.
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane, and we therefore speculate that this star-formation event was triggered by the last disk midplane passage of the Magellanic stream gas, which occurred at a Galactocentric radius $R \approx 30~\kpc$.
However, most details of this idea remain a mystery: the Magellanic stream is low column density, the Milky Way disk at this large radius has low surface density, and the relative velocity of the stream gas and MW gas is large.
However it formed, the discovery of a young stellar cluster in the Milky Way halo presents an interesting opportunity for study.
\end{abstract}

\keywords{Galaxy: halo}

@@

@@

% Notebook: Figure-CMDs
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/gaia-cmd-pm.pdf}
\caption{Color-magnitude diagram, sky positions, and proper motions from \gaia\ \DR{2} for the region around $(\alpha, \delta) \sim (179, -29)^\circ$.
\textbf{Left panel}: \gaia\ color-magnitude diagram (CMD) for all sources with $G < 20$ and sky position within 3 degrees of $(\alpha, \delta) = (179.5, -28.8)^\circ$.
Red line shows a $100~\textrm{Myr}$, $\feh = -1.1$ \acronym{MIST} isochrone shifted to a distance of $30~\kpc$.
Shaded pixels (orange) show regions of high density and points (black markers) are plotted when there are fewer than 8 stars per pixel.
\textbf{Middle panel}: Sky positions for all sources in the shown sky region (grey 2D histogram).
Black markers show only sources in the blue selection polygon shown in the CMD (left panel).
\textbf{Right panel}: Proper motions for all sources in the shown sky region (grey 2D histogram).
Black markers again show only sources in the blue CMD selection region (left panel).
}
\label{fig:cmds}
\end{figure}

@@

 % Notebook: Figure-DECam-CMD
 \begin{figure}
 \centering
 \includegraphics[width=0.6\textwidth]{figures/DECam-field.pdf}
 \caption{The sky region around \clustername, showing the same \gaia\ CMD-selected sources as in \figurename~\ref{fig:cmds} (black markers), and point sources identified from \decam\ $g$-band follow up (blue and red markers).
 Red markers show sources in the \decam\ field used as control sources, and blue markers show sources used as cluster member candidates.}
 \label{fig:decam-field}
 \end{figure}

% Notebook: Figure-DECam-CMD
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{figures/DECam-cmd.pdf}
\caption{\decam\ $g-i$ versus $g$ color-magnitude diagrams for the \decam\ control fields (left panel) and \decam\ cluster fields (right panel).
Note the prominent young, distant main sequence in the cluster fields: This is the main sequence of \clustername.
The sources in the dashed (red) rectangle in the right panel are later used to measure the cluster stellar population parameters.
}
\label{fig:decam-cmd}
\end{figure}

@@

@@

% Notebook: Proper-motion-membership
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/pm-model.pdf}
\caption{\textbf{Left panel}: Sky positions of all \gaia\ sources in this region with $\bprp < 0.35$ (grey markers).
The circles show the regions used as cluster and control fields in the proper motion inference (\sectionname~\ref{sec:pmclean}).
The control fields, together, have the same area as the cluster field.
\textbf{Right panel}: The colored 2D density shows the difference in number of sources per pixel in the cluster field versus the two control fields (see left panel).
The clear over-density in the cluster field (blue) are stars in \clustername.}
\label{fig:pm-members}
\end{figure}

@@

% Notebook: Figure-hierarch-results
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{figures/hierarch-corner.pdf}
\caption{A corner plot showing marginal posterior probability distributions estimated from the the posterior samples generated from the hierarchical inference of the cluster population parameters.}
\label{fig:hierarch-corner}
\end{figure}

@@

% Notebook: Figure-hierarch-results
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{figures/hierarch-results.pdf}
\caption{The \decam\ color-magnitude diagram for sources with probabilty $>0.5$ of belonging to \clustername\ (markers).
The solid line (blue) shows the \acronym{MIST} isochrone for the median posterior sample from the hierarchical inference of the cluster population parameters, shifted to the inferred distance to \clustername, and extincted with the inferred extinction.
The dashed line (green) shows the same isochrone, shifted $\approx 0.75$ magnitudes brighter, representing the expected location of the equal-mass binary sequence of the cluster.
}
\label{fig:hierarch-iso}
\end{figure}

>>>>

% TODO
% - Sand, Seth, et al. Virgo: blue stars with nothing else
% - Split cluster into 2? Price-Whelan 1a and 1b?

% \documentclass[modern]{aastex62}
\documentclass[twocolumn]{aastex62}
\usepackage{amsmath}
\graphicspath{{./}{figures/}}

@@

We report the discovery of a young $\tau \sim \clage$, low-mass $M \sim 1200~\msun$, metal-poor $[\textrm{Fe}/\textrm{H}] \sim \clfeh$ stellar association in the Milky Way halo, \clustername.
At a Galactocentric (cylindrical) position $(R, z) \sim (23, 15)~\textrm{kpc}$, the association is (on the sky) near the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the Magellanic stream, $D \approx \cldist$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way that do not account for ram pressure and gas interaction with MW disk.
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane, and we therefore speculate that this star-formation event was triggered by the last disk midplane passage of the Magellanic stream gas, which occurred at a Galactocentric radius $R \approx 30~\kpc$.
However, most details of this idea remain a mystery: the Magellanic stream is low column density, the Milky Way disk at this large radius has low surface density, and the relative velocity of the stream gas and MW gas is large.
However it formed, the discovery of a young stellar cluster in the Milky Way halo presents an interesting opportunity for study.
\end{abstract}

\keywords{
    Galaxy: open clusters and associations --
    Galaxy: halo --
    stars: formation --
    surveys}

@@

% Notebook: Figure-CMDs
\begin{figure*}[htb!]
\centering
\includegraphics[width=\textwidth]{figures/gaia-cmd-pm.pdf}
\caption{Color-magnitude diagram, sky positions, and proper motions from \gaia\ \DR{2} for the region around $(\alpha, \delta) \sim (179, -29)^\circ$.
\textbf{Left panel}: \gaia\ color-magnitude diagram (CMD) for all sources with $G < 20$ and sky position within 3 degrees of $(\alpha, \delta) = (179.5, -28.8)^\circ$.
Red line shows a $100~\textrm{Myr}$, $\feh = -1.1$ \acronym{MIST} isochrone shifted to a distance of $30~\kpc$.
Shaded pixels (orange) show regions of high density and points (black markers) are plotted when there are fewer than 8 stars per pixel.
\textbf{Middle panel}: Sky positions for all sources in the shown sky region (grey 2D histogram).
Black markers show only sources in the blue selection polygon shown in the CMD (left panel).
\textbf{Right panel}: Proper motions for all sources in the shown sky region (grey 2D histogram).
Black markers again show only sources in the blue CMD selection region (left panel).
}
\label{fig:cmds}
\end{figure*}

@@

@@

 % Notebook: Figure-DECam-CMD
 \begin{figure}
 \centering
 \includegraphics[width=0.45\textwidth]{figures/DECam-field.pdf}
 \caption{The sky region around \clustername, showing the same \gaia\ CMD-selected sources as in \figurename~\ref{fig:cmds} (black markers), and point sources identified from \decam\ $g$-band follow up (blue and red markers).
 Red markers show sources in the \decam\ field used as control sources, and blue markers show sources used as cluster member candidates.}
 \label{fig:decam-field}
 \end{figure}

% Notebook: Figure-DECam-CMD
\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/DECam-cmd.pdf}
\caption{\decam\ $g-i$ versus $g$ color-magnitude diagrams for the \decam\ control fields (left panel) and \decam\ cluster fields (right panel).
Note the prominent young, distant main sequence in the cluster fields: This is the main sequence of \clustername.
The sources in the dashed (red) rectangle in the right panel are later used to measure the cluster stellar population parameters.
}
\label{fig:decam-cmd}
\end{figure}

@@

% Notebook: Proper-motion-membership
\begin{figure*}[ht!]
\centering
\includegraphics[width=\textwidth]{figures/pm-model.pdf}
\caption{\textbf{Left panel}: Sky positions of all \gaia\ sources in this region with $\bprp < 0.35$ (grey markers).
The circles show the regions used as cluster and control fields in the proper motion inference (\sectionname~\ref{sec:pmclean}).
The control fields, together, have the same area as the cluster field.
\textbf{Right panel}: The colored 2D density shows the difference in number of sources per pixel in the cluster field versus the two control fields (see left panel).
The clear over-density in the cluster field (blue) are stars in \clustername.}
\label{fig:pm-members}
\end{figure*}

@@

@@

% Notebook: Figure-hierarch-results
\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{figures/hierarch-corner.pdf}
\caption{A corner plot showing marginal posterior probability distributions estimated from the the posterior samples generated from the hierarchical inference of the cluster population parameters.}
\label{fig:hierarch-corner}
\end{figure}

@@

% Notebook: Figure-hierarch-results
\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{figures/hierarch-results.pdf}
\caption{The \decam\ color-magnitude diagram for sources with probabilty $>0.5$ of belonging to \clustername\ (markers).
The solid line (blue) shows the \acronym{MIST} isochrone for the median posterior sample from the hierarchical inference of the cluster population parameters, shifted to the inferred distance to \clustername, and extincted with the inferred extinction.
The dashed line (green) shows the same isochrone, shifted $\approx 0.75$ magnitudes brighter, representing the expected location of the equal-mass binary sequence of the cluster.
}
\label{fig:hierarch-iso}
\end{figure}
